MapReduce to My Ears\\ {\Large Using Big Data for Music Synthesis}

### 2013-05-10 writeup/paper.tex
adding amanda's changes

\begin{abstract}
\end{abstract}

@@

The big data revolution has revolutionized the way we think about,
design, and implement algorithms at scale.  As such, problems from all
different areas of computer science have been lifted into the
MapReduce framework.  By contrast, MapReduce has been used relatively
less to explore art or creative tasks.  Our project looks to answer:
what kinds of art can we create harnessing large amounts of data at
scale?

@@

The core idea in converting image data to music is to view the
information in an image (color, size, opacity, etc..) as a mapping to
notes and durations with which those notes are sustained.  In our
system we chose MIDI as the output format.  Our system can then be
modeled as a function from image data to MIDI output.  Note that the
choice of MIDI limits the expressivity of our music.  Going to other
formats (such as MP3 or OOG) could provide a larger range of
primitives with which to construct music, but does not enjoy the
simplicity that MIDI provides.

@@

As a result, we produced various pieces of music.  Our longest piece
(named \code{Digital Nightmare}) was over six minutes long and
composed of over 3000 images.  We also produced shorter (self
contained) pieces from various related images: such as GIFs of cat
jumping animations.  All of these pieces are made available in our
repository.

@@

Many domain specific programming languages exist for music synthesis
[\cite{haskore},\cite{lilypad}].  Many of these languages synthesize
music in the small ().  By contrast, our approach focuses on using
large amounts of data to create music...

>>>>

\begin{abstract}

\end{abstract}

@@

The big data revolution has revolutionized the way we think about,
design, and implement algorithms at scale.  As such, problems from all
different areas of computer science have been lifted into the
MapReduce framework.  While MapReduce has been applied to many
large-scale data processing domains, it has been used relatively less
to explore art or creative tasks.  Our project looks to answer: what
kinds of art can we create harnessing large amounts of data at scale?

@@

The core idea in converting image data to music is to view the
information in an image (color, size, opacity, etc..) as a mapping to
notes and durations with which those notes are sustained.  In our
system we chose MIDI as the output format.  Our system can then be
modeled as a function from image data to MIDI output.  Using MIDI as a
format works well because it is easy to work with (we can model music
as MIDI events, which are simply streams of musical data) and can be
mixed by many synthesizers to produce music using a variety of
instruments.

@@

As a result, we produced various pieces of music.  Our longest piece
(named \code{Digital Nightmare}) was over six minutes long and
composed of over 3500 images.  We also produced shorter (self
contained) pieces from various related images: such as GIFs of cat
jumping animations.  All of these pieces are made available in our
repository.  We ran the entire MapReduce job on the Hoth cluster to
produce our longest piece.

@@

Our work is inspired by the music visualization effects included in
media player software (which traditionally includes effects such as
oscilloscopes and frequency representations).  However, many domain
specific programming languages exist for music synthesis.  In
particular we looked at Haskore \cite{haskore} and
LilyPond\cite{lilypond}.  Haskore is a language based on Haskell,
which allows the user to express mathematically beautiful music using
an algebraic composition strategy.  LilyPond is a specification
language used to create high quality sheet music (similar to \LaTeX\ 
for music).  Many of these languages synthesize music in the small,
using a human based specification language.  By contrast, our approach
focuses on using large amounts of data to create music: rather than a
human specification.  We are unaware of any projects that use big data
to produce art or other forms of entertainment.
